---
title: "Using ggerr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using_ggerr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggerr)
library(ggplot2)
```

The `ggerr` package makes it easier to add error-bars to pplots.

Adding errorbars of +/- one standard deviation in each group is done like this:

```{r}
ggplot(iris, aes(x = Petal.Length, y = Petal.Width, color = Species)) + geom_point() + stat_err() + stat_err(geom="point")
```
The second call to `stat_err()` with geom "point" is to add the point at the mean. 

The `stat_err()` detaults to the geom "errorbar", so the width of the whiskers are controlled through that:

```{r}
ggplot(iris, aes(x = Petal.Length, y = Petal.Width, color = Species)) + geom_point() + stat_err(width=.1) + stat_err(geom="point")
```

Horizontal errorbars are obtained by the geom errorbarh. 
The `stat_err()` knows about standard deviation and standard error of the mean. They are selected by the `spread` parameter. It is possible to multiply the errorbars by a factor, using the `mult` parameter.

The grouping does not need to be coltrolled by the `color` aestetics, it can also be done through an explicit `group` aestetics.

As an example, we add vertical and horizontal errorbars of two times the standard error of the mean:

```{r}
ggplot(iris, aes(x = Petal.Length, y = Petal.Width, group = Species)) + geom_point() + stat_err(spread = "se", mult = 2, width=.1) + stat_err(geom="point") + stat_err(spread = "se", mult = 2, geom = "errorbarh", height = .1)
```

